Tracing the Hercules stream around the Galaxy

### 2010-05-28 tex/hercules.tex
groups-simulations: arXiv reference for Shen paper
groups-simulations: small tweaks

\begin{abstract}
It has been proposed that the Hercules stream, a group of co-moving
stars in the Solar neighborhood offset from the bulk of the velocity
distribution, is the result of resonant interactions between stars in
the outer disk and the Galactic bar. So far it has only been seen in
the immediate Solar neighborhood, but any resonance model make a
prediction over a large fraction of the Galactic disk. I predict the
distribution of stellar velocities and changing Hercules signature as
a function of location in the Galactic disk in a simple model for the
Galaxy and the bar that produces the observed Hercules stream. The
Hercules feature is expected to be strong enough to be unambiguously
detected in the distribution of line-of-sight velocities in selected
directions. I identify quantitatively the most promising lines of
sight for detection in line-of-sight velocities using the
Kullback-Leibler divergence between the predictions from the resonance
model and an axisymmetric model; these directions are at $250^{\circ}
\lesssim l \lesssim 290^{\circ}$. The predictions presented here are
only weakly affected by distance uncertainties, assumptions about the
distribution function in the stellar disk, and the details of the
Galactic potential including the effect of spiral structure. \Gaia\
and future spectroscopic surveys of the Galactic disk such as \apogee\
will be able to robustly test the origin of the Hercules stream and
constrain the properties of the Galactic bar.
\end{abstract}

@@

Rather than wait for a comprehensive survey of the Galactic disk that
measures the full phase-space position of a large number of stars by
measuring proper motions, we can ask whether a targeted survey of
line-of-sight velocities in selected regions of the Galaxy could
detect a clear signature of the Hercules feature? The Hercules feature
stands out in the velocity distribution most in its tangential
velocity $\vphi$. On the Solar circle near the Sun the line-of-sight
velocity is almost entirely this Galactocentric tangential velocity,
so we would expect to see at least a hint of the Hercules stream in
the line-of sight.

In \figurename~\ref{fig:rphi1d} I show the predicted line-of-sight
velocity distributions in various locations near the Solar circle all
around the Galaxy, obtained by integrating the full two-dimensional
velocity distribution over the velocity perpendicular to the line of
sight. This \figurename\ is no longer symmetric with respect to shifts
of 180$^{\circ}$ degree in azimuth, since the line-of-sight direction
breaks this symmetry. The gray curve is the line-of-sight distribution
we would observe if the targeted stars were in a steady state in an
axisymmetric potential with a Dehnen distribution function
(\eqnname~[\ref{eq:fdehnen}]). Both line-of-sight velocity
distributions are normalized to have equal mass since I assume that
our knowledge of the Galaxy and the tracer population will not be
great enough that we can test predicted differences in the absolute
number of stars at a given location between the axisymmetric and the
bar-resonance predictions. These differences are also largely a
function of the particular form and parameters of the distribution
function that we choose to represent the stellar disk before bar
formation.

@@

\bibitem[Shen \etal(2010)]{Shen10a}
  Shen,~J., \etal, 2010, \apj L, submitted

>>>>

\begin{abstract}
It has been proposed that the Hercules stream, a group of co-moving
stars in the Solar neighborhood offset from the bulk of the velocity
distribution, is the result of resonant interactions between stars in
the outer disk and the Galactic bar. So far it has only been seen in
the immediate Solar neighborhood, but the resonance model makes a
prediction over a large fraction of the Galactic disk. I predict the
distribution of stellar velocities and the changing Hercules feature
in this distribution as a function of location in the Galactic disk in
a simple model for the Galaxy and the bar that produces the observed
Hercules stream. The Hercules feature is expected to be strong enough
to be unambiguously detected in the distribution of line-of-sight
velocities in selected directions. I identify quantitatively the most
promising lines of sight for detection in line-of-sight velocities
using the Kullback-Leibler divergence between the predictions of the
resonance model and an axisymmetric model; these directions are at
$250^{\circ} \lesssim l \lesssim 290^{\circ}$. The predictions
presented here are only weakly affected by distance uncertainties,
assumptions about the distribution function in the stellar disk, and
the details of the Galactic potential including the effect of spiral
structure. \Gaia\ and future spectroscopic surveys of the Galactic
disk such as \apogee\ will be able to robustly test the origin of the
Hercules stream and constrain the properties of the Galactic bar.
\end{abstract}

@@

Rather than wait for a comprehensive survey of the Galactic disk that
measures the full phase-space position of a large number of stars by
measuring proper motions, we can ask whether a targeted survey of
line-of-sight velocities in selected regions of the Galaxy could
detect a clear signature of the Hercules feature? The Hercules feature
stands out in the velocity distribution most in its tangential
velocity. On the Solar circle near the Sun the line-of-sight velocity
is almost entirely this Galactocentric tangential velocity, so we
would expect to see at least a hint of the Hercules stream in the
line-of sight.

In \figurename~\ref{fig:rphi1d} I show the predicted line-of-sight
velocity distributions in various locations near the Solar circle all
around the Galaxy, obtained by integrating the full two-dimensional
velocity distribution over the velocity component perpendicular to the
line of sight. This \figurename\ is no longer symmetric with respect
to shifts of 180$^{\circ}$ degree in azimuth, since the line-of-sight
direction breaks this symmetry. The gray curve is the line-of-sight
distribution we would observe if the targeted stars were in a steady
state in an axisymmetric potential with a Dehnen distribution function
(\eqnname~[\ref{eq:fdehnen}]). Both line-of-sight velocity
distributions are normalized to have equal mass since I assume that
our knowledge of the Galaxy and the tracer population will not be good
enough that we can test predicted differences in the absolute number
of stars at a given location between the axisymmetric and the
bar-resonance predictions. These differences are also largely a
function of the particular form and parameters of the distribution
function that we choose to represent the stellar disk before bar
formation.

@@

\bibitem[Shen \etal(2010)]{Shen10a}
  Shen,~J., \etal, 2010, \apj L, submitted, arXiv:1005.0385v1 [astro-ph.CO]
